---
title: 'Appendices: Covariate tests'
output:
  pdf_document: default
header-includes: \usepackage[labelformat=empty]{caption}
---

```{r child = file.path(here::here(),'setup.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Functions.Rmd')}
```

```{r child = file.path(here::here(),'AppendixFiles','Appendix-Variables.Rmd')}
```


```{r echo=FALSE}
addchl = TRUE
add1956 = FALSE
```


```{r spawncovnames, echo=FALSE}
vars <- getspawnercovs()
datcommon <- dat.spawners[, 
  c("Year", "spawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", vars$allcovnames)]
dat <- na.omit(datcommon)

NE=mean(abs(dat$spawners0-dat$spawners1))
```

```{r spawnerapptabeqns, echo=FALSE}
nullmod <- "$ln(S_t) = ln(S_{t-1}) + \\epsilon_t$"
basemod <- "$ln(S_t) = \\alpha + s(ln(W_{t-1})) + \\epsilon_t$"
resp <- "ln(S_t)"
nullformula <- "spawners0 ~ -1 + offset(spawners1)"
baseformula <- "spawners0 ~ s(nspawners1, sp=0.6)"
```

```{r spawntable}
seltable <- makeapptable(dat, nullmod, basemod, nullformula, baseformula, resp, vars$covnamelist, vars$covlags, include.num = TRUE)
```

```{r spawnchlnames, echo=FALSE, eval=addchl}
vars <- getspawnerchlcovs()
datcommon <- dat.chl[, 
  c("Year", "spawners0", "spawners1", "spawners2",
    "nspawners1", "nspawners2", vars$allcovnames)]
dat <- na.omit(datcommon)
NE=mean(abs(dat$spawners0-dat$spawners1))
```

```{r tablechlvals, eval=addchl}
basemod <- "$ln(S_t) = \\alpha + p(ln(W_{t-1})) + \\epsilon_t$"
baseformula <- "spawners0 ~ poly(nspawners1, 2, raw=TRUE)"
m3form <- "poly(cov, 2, raw=TRUE)"
```

```{r chltable, eval=addchl}
seltable = rbind(seltable, spacetmp)
chlbody <- makeapptable(dat, nullmod, basemod, nullformula, baseformula, resp, vars$covnamelist, vars$covlags, m3form=m3form, include.num = TRUE)
seltable <- rbind(seltable, chlbody)
```

```{r print-apptable1, echo=FALSE}
seltable = seltable[,c("Model", "Residual.df", "Adj.R2", "RMSE", "AIC", "LOOCV", "LOOMd")]
colnames(seltable) <- c("Model", "Resid. df", "Adj. R2", "RMSE", "AICc", "LOOCV", "LOOMdAE")
if(isLatex) colnames(seltable)=c("Model", "\\shortstack{Resid.\\\\df}", "\\shortstack{Adj.\\\\$R^2$}", "RMSE", "AICc", "\\shortstack{LOOCV\\\\RMSE}", "\\shortstack{LOOCV\\\\MdAE}")

thecap=paste("Covariate tests for the Jul-Sep catch ($S_t$). M is the base model with only prior season Oct-Mar catch ($W_{t-1}$) as the covariate. To the base model, the environmental covariates are added. Nearshore is 0-80km and regional is 0-160km. The SST data are from the Optimum Interpolation (ioSST) AVHRR data set. The models are nested sets, e.g. 1, 2a, 3a and 1, 2b, 3b.")
fullcap=paste("Table ", ref("apptab:spawnercovariate-models", pre="A"), ". ", thecap, sep="")

if(isLatex){
  latexkable(seltable, fullcap)
}else{
 kable(seltable, align=paste0("l", paste0(rep("c",dim(seltable)[2]-1),collapse="")), caption=fullcap, escape=FALSE)
}


```

Notes: LOOCV = Leave one out cross-validation. RMSE = root mean square error. MdAE = median absolute error. AICc = Akaike Information Criterion corrected for small sample size. $\dagger$ and $\dagger\dagger$ = AICc greater than 2 and greater than 5 below model M (base catch model).  $\ddagger$, $\ddagger\ddagger$, and ${\ddagger}{\ddagger}{\ddagger}$  = LOOCV RMSE 5\%, 10\% and 20\% below model M, respectively. $t$ indicates current year and $t-1$ is the prior year. $W_t$ spans two calendar years (Oct-Mar); $t$ is the year in Oct. Thus if $t=2014$, $S_t$ is Jul 2014 to Sep 2014 and $W_{t-1}$ is Oct 2013 to Mar 2014. For covariates that are multiyear, such as the multiyear average SST, $t$ is the calendar year at the end of the multiyear span; thus the 2.5 year average SST for 2014 is Jan 2012 to Jun 2014.