---
title: 'Supplemental Information: Full model tests and diagnostics'
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: \usepackage[labelformat=empty]{caption}
---

```{r child = file.path(here::here(), 'setup.Rmd')}

```

```{r echo=FALSE}
Supplement <- TRUE # Change if this will be labelled appendix
if (!exists(".rmdenvir")) .rmdenvir <- environment()
```

## Tests for prior season catch as covariate

```{r child = file.path(here::here(), 'SupplementFiles', 'Table App A spawner model.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = file.path(here::here(), 'SupplementFiles', 'Table App A nonspawner model.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## Tests for environmental variables as covariates


```{r child = file.path(here::here(), 'AppendixFiles', 'AppendixNonSpawner.Rmd')}
```


## Tests with upwelling and precipitation interactions

```{r child = file.path(here::here(), 'SupplementFiles', 'Table_Precip_Interactions.Rmd')}
```


```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## Validation of catch base models

```{r child = file.path(here::here(), 'SupplementFiles', 'Table App D.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## Comparison of land and oceanic rainfall measurements

```{r echo=FALSE}
# set the table pre
if (Supplement) {
  pre <- "S"
  figset <- "figsupp:"
} else {
  pre <- "E"
  figset <- "figappE:"
}
```

```{r echo=FALSE}
thecap <- paste0("Monthly precipitation measured over land via land gauges versus the precipitation measured via remote sensing over the ocean.")
fullcap.precip <- paste0("Figure ", ref(paste0(figset, "figprecip"), pre = pre), ". ", thecap)
```

```{r rainfall.mon.comparison, echo=FALSE,fig.cap=fullcap.precip, fig.width=6,fig.height=6,message=FALSE,warning=FALSE}
plot(seio_covariates_mon$Precip.Kerala, seio_covariates_mon$precip.gpcp.kerala,
  ylab = "Monthly land precipitation (mm)",
  xlab = "Monthly ocean precipitation (average mm/day)"
)
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## Comparison of multiyear average regional SST from AVHRR and ICOADS

```{r echo=FALSE}
thecap <- paste0("Multiyear average regional SST from the AVHRR versus ICOADS data sets.")
fullcap.dmi <- paste0("Figure ", ref(paste0(figset, "figsstvdmi"), pre = pre), ". ", thecap)
```

```{r sst.dmi.comparison, echo=FALSE,fig.cap=fullcap.dmi, fig.width=6,fig.height=6,message=FALSE,warning=FALSE}
plot(fullrespdat$SSTICOAD.2.10.3.yr.runsum.0, fullrespdat$SST.2.10.3.yr.runsum.0, xlab = "2.5-year average SST - ICOADS", ylab = "2.5-year average regional SST - AVHRR")
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## SST Product Comparisons

```{r child = file.path(here::here(), 'SupplementFiles', 'SST_Comparisons.Rmd')}
```


```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

## Comparison of multiyear average regional SST and ocean climate indices

```{r echo=FALSE}
thecap <- paste0("Multiyear average regional SST from AVHRR versus ocean climate indices (AMO, ONI, PDO and DMI) 1983-2016. The corvariates are those used in the analyses. Multi-year SST is the 2.5-year average of the regional (0-160km from coast) SST, so January $t-2$ to July $t$. The climate indices are 12-month average from July $t-1$ to June $t$. For the analyses, Sep-Nov average DMI in the prior year was used so that is also added.")
fullcap.sstvamo <- paste0("Figure ", ref(paste0(figset, "figsstvamo"), pre = pre), ". ", thecap)
```

```{r sst.amo.comparison, echo=FALSE,fig.cap=fullcap.sstvamo, fig.width=6, fig.height=6, message=FALSE, warning=FALSE}
#library("ggplot2")
#library("GGally")
dat <- fullrespdat
df <- data.frame(SST = scale(dat$SST.2.10.3.yr.runsum.0), AMO = scale(dat$AMO.mon7to6.0), PDO = scale(dat$PDO.mon7to6.0), ONI = scale(dat$ONI.mon7to6.0), DMI = scale(dat$DMI.mon7to6.0), DMI2 = scale(dat$DMI.mon9to11.1))
GGally::ggpairs(df, columnLabels = c("2.5-yr SST", "Jul-Jun AMO", "Jul-Jun PDO", "Jul-Jun ONI", "Jul-Jun DMI", "Sep-Nov DMI"))
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```
