---
title: 'Supplemental Information: Full model tests and diagnostics'
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: \usepackage[labelformat=empty]{caption}
---

```{r child = 'setup.Rmd'}
```

```{r echo=FALSE}
Supplement <- TRUE # Change if this will be labelled appendix
if(!exists(".rmdenvir")) .rmdenvir = environment()
# here::set_here()
```

# Tests for prior season catch as covariate

```{r child = file.path('SupplementFiles', 'Table App A spawner model.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

```{r child = file.path('SupplementFiles', 'Table App A nonspawner model.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Tests for environmental variables as covariates

```{r child = file.path('SupplementFiles', 'Table App B spawner covariates.Rmd'), eval=FALSE}
```

```{r, results="asis", echo=FALSE, eval=FALSE}
cat("\\clearpage")
```

```{r child = file.path('SupplementFiles', 'Table App B nonspawner covariates M with S2.Rmd'), eval=FALSE}
```

```{r, results="asis", echo=FALSE, eval=FALSE}
cat("\\clearpage")
```

```{r child = file.path('SupplementFiles', 'Table App B nonspawner covariates M wo S2.Rmd'), eval=FALSE}
```

```{r child = file.path('AppendixFiles', 'AppendixNonSpawner.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

<!--
# Tests for Chlorophyll-a as a covariate
-->

```{r child = file.path('SupplementFiles', 'Table App C chlorophyll.Rmd'), eval=FALSE}
```

```{r, results="asis", echo=FALSE, eval=FALSE}
cat("\\clearpage")
```

# Tests with upwelling and precipitation interactions

```{r child = file.path('SupplementFiles', 'Table_Precip_Interactions.Rmd')}
```


```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Validation of catch base models

```{r child = file.path('SupplementFiles', 'Table App D.Rmd')}
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Comparison of land and oceanic rainfall measurements

```{r echo=FALSE}
#set the table pre
if(Supplement){
  pre <- "S"
  figset <- "figsupp:"
}else{
  pre <- "E"
  figset <- "figappE:"
}
```

```{r echo=FALSE}
thecap=paste0("Monthly precipitation measured over land via land gauges versus the precipitation measured via remote sensing over the ocean.")
fullcap.precip <- paste0("Figure ", ref(paste0(figset,"figprecip"), pre=pre), ". ", thecap)
```

```{r rainfall.mon.comparison, echo=FALSE,fig.cap=fullcap.precip, fig.width=6,fig.height=6,message=FALSE,warning=FALSE}
plot(seio_covariates_mon$Precip.Kerala, seio_covariates_mon$precip.gpcp.kerala, ylab="Monthly land precipitation (mm)",
xlab="Monthly ocean precipitation (average mm/day)")
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Comparison of multiyear average regional SST from AVHRR and ICOADS

```{r echo=FALSE}
thecap=paste0("Multiyear average regional SST from the AVHRR versus ICOADS data sets.")
fullcap.dmi <- paste0("Figure ", ref(paste0(figset,"figsstvdmi"), pre=pre), ". ", thecap)
```

```{r sst.dmi.comparison, echo=FALSE,fig.cap=fullcap.dmi, fig.width=6,fig.height=6,message=FALSE,warning=FALSE}
plot(fullrespdat$SSTICOAD.2.10.3.yr.runsum.0, fullrespdat$SST.2.10.3.yr.runsum.0, xlab="2.5-year average SST - ICOADS", ylab="2.5-year average regional SST - AVHRR")
```

```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# SST Product Comparisons

```{r child = file.path('SupplementFiles', 'SST_Comparisons.Rmd')}
```


```{r, results="asis", echo=FALSE}
cat("\\clearpage")
```

# Comparison of multiyear average regional SST and ocean climate indices

```{r echo=FALSE}
thecap=paste0("Multiyear average regional SST from AVHRR versus ocean climate indices (AMO, ONI, PDO and DMI) 1983-2016. The corvariates are those used in the analyses. Multi-year SST is the 2.5-year average of the regional (0-160km from coast) SST, so January $t-2$ to July $t$. The climate indices are 12-month average from July $t-1$ to June $t$. For the analyses, Sep-Nov average DMI in the prior year was used so that is also added.")
fullcap.sstvamo <- paste0("Figure ", ref(paste0(figset,"figsstvamo"), pre=pre), ". ", thecap)
```

```{r sst.amo.comparison, echo=FALSE,fig.cap=fullcap.sstvamo, fig.width=6,fig.height=6,message=FALSE,warning=FALSE}
library("ggplot2")                     # Load ggplot2 package
library("GGally")    
dat <- fullrespdat
df <- data.frame(SST=scale(dat$SST.2.10.3.yr.runsum.0), AMO=scale(dat$AMO.mon7to6.0), PDO=scale(dat$PDO.mon7to6.0), ONI=scale(dat$ONI.mon7to6.0), DMI=scale(dat$DMI.mon7to6.0), DMI2=scale(dat$DMI.mon9to11.1))
ggpairs(df, columnLabels=c("2.5-yr SST", "Jul-Jun AMO", "Jul-Jun PDO", "Jul-Jun ONI", "Jul-Jun DMI", "Sep-Nov DMI"))
```



